ActiVis: Mobile Object Detection and Active Guidance for People with Visual Impairments

### 2019-07-05 paper.tex

\begin{abstract}
  The ActiVis project aims to deliver a mobile system that is able to guide a person with visual impairments towards a target object or area in an unknown indoor environment. 
  For this, it uses new developments in object detection, mobile computing, action generation and human-computer interfacing to interpret the user's surroundings and present effective guidance directions.
  Our approach to direction generation uses a Markov Decision Process to track the system's state and output the optimal location to be investigated.
  This system includes an object detector and an audio-based guidance interface to provide a complete active search pipeline.
  The ActiVis system was evaluated in a set of experiments showing better performance than a simpler unguided case.
  \keywords{Active vision, vision impairment, object detection}
\end{abstract}

@@

The issue with a marker-based approach is that it requires significant effort to place and maintain them in an environment, which is remedied by markerless systems.
However, both of these methods use passive guidance approaches that rely on the user placing the desired marker/object within the camera's view by themselves before any guidance is provided. 
VizWiz~\cite{bigham2010vizwiz} can leverage a human's understanding of the environment to guide a user to the correct location, but there is significant lag and a reliance on a good internet connection and remote worker being available. 
Previous work on the ActiVis system~\cite{lock2019active} addressed the passive guidance issue by implementing an MDP-based controller that gives the user pointing instructions to find an out-of-view object, showing that this is a viable method for object search. 
However, that work used QR-codes to simulate real objects and on-screen prompts to present the guidance instructions.
In this paper we expand the control model, replacing the QR-codes with real objects and providing audio guidance, thereby implementing a complete object detection and guidance pipeline for people with vision impairments.

@@

A complete system diagram for ActiVis is given in Fig.~\ref{fig:sys-diagram}. 
This diagram shows a typical feedback control loop that generates a control signal to minimise some error signal, \emph{e}, and drive the output to some reference, $r$, but has been modified to incorporate a human within the loop.
In this case, $r$ is the target object.

@@

Adding a human into the loop adds an additional block, \emph{H}, representing a human that receives a control signal, \emph{u}, from the controller, \emph{K}. 
However, the challenge is that a person may interpret \emph{u} in any unpredictable way that results in the signal \emph{u*} that points the camera, \emph{P}, to a new location containing a new object observation, \emph{y}.
It is therefore important to design \emph{K} to be robust enough to accommodate different user habits and limitations and ensure that \emph{u*} tracks \emph{u} as closely as possible. 
The object detector's classification errors, \emph{n}, is added to the feedback loop as a noise signal and affects \emph{K}'s output.

In our previous work~\cite{lock2019active}, we assumed a perfect classifier and solved the problem using a Markov decision process (MDP). 
In this work we do away with that somewhat naive assumption and replace the MDP controller with a POMDP-based controller that can handle classification uncertainties in object detectors and classifiers. 
This controller works by generating a trail of virtual waypoints for the user to point the camera towards that eventually leads to the target object.
The waypoint positions are based on the model's pre-trained internal knowledge of the inter-object spatial relationships and they are placed in a way that maximises the probability of the user finding the target object.
The design and implementation of this controller are discussed next.

@@

A POMDP is an extension to the MDP model that allows it to handle cases where the state is not directly observable, allowing it to be used in more realistic scenarios. 
The implication of this is that a POMDP-based agent does not know its state at any point in time, but must infer it based on the known model parameters and sensor accuracy, which is the object detector's accuracy in this case.
This state inference relies on a so-called belief meta-state that is updated with additional observations to reflect the likelihood that the agent is in any given state.
The belief state is fully observable by the agent and can be used to infer the mobile device's current state and generate the next action.

A POMDP model is represented by the 8-tuple $(\mathbf{S}, \mathbf{A}, \mathbf{T}, \mathbf{R}, \mathbf{\Omega}, \mathbf{O}, \mathbf{b}, \gamma)$, where $\mathbf{S}$ represents a finite set of discrete states, $\mathbf{A}$ is a set of discrete actions, $\mathbf{T}$ is a matrix containing the probabilities of transitioning from state $s$ to state $s'$, $s, s' \in \mathbf{S}$, after executing action $a$, $a \in \mathbf{A}$, and $\mathbf{R}$ is the reward the agent receives executing action $a$ and reaching state $s'$.
$\mathbf{\Omega}$ is the set of possible state observations, while the matrix $\mathbf{O}$ contains the probabilities of making state observation $\omega$, $\omega \in \mathbf{\Omega}$, when in state $s$ after executing action $a$.
Finally, $\mathbf{b}$ is the belief vector containing the state probability distribution and $\gamma$ is a discount factor that prioritises long-term over short-term rewards and affects the model's convergence rate. 
The POMDP's parameters and training process are discussed next.

@@

The state is given by $s = \langle u, n, v \rangle$, where $u, u\in\mathbf{U}$, is the object within view, $n$ the number of search steps taken during search and a binary variable $v$ that tracks the uniqueness of a waypoint's position for the search.

The possible actions that dictates the location the mobile device will generate the next waypoint are given by $\mathbf{A} = \{ \textrm{UP, DOWN, LEFT, RIGHT} \}$.

$\mathbf{T}$ was determined by extracting the inter-object spatial relationships for a limited number of objects in terms of the actions in $\mathbf{A}$ from the OpenImages~\cite{openimages} dataset. 
For example, by iterating over the images containing the objects of interest, we can see that the object `monitor' is located above (i.e. UP) the object `keyboard' in 16\% of images containing both objects (see Fig.~\ref{fig:openimage}). 
The transition function for this case is $t(s, a, s') = t(\textrm{keyboard}, \textrm{UP}, \textrm{monitor}) = 0.16$.

@@

The reward function encourages the device controller to search for and find the target object by giving it a substantial reward for doing so, while penalising it for every action that does not result in it finding the target object.
Furthermore, additional penalties are given if the controller generates a waypoint in an area it has explored before $(v = \textrm{true}$) or when it exceeds some search length threshold denoted by $n_{\max}$.
The reward values were selected based on lessons learned during the implementation of the MDP model and the rewards for both that and this work are summarised in Table~\ref{tab:rewards}.

\begin{table}
  \centering
  \caption{The reward function for the POMDP. }\label{tab:rewards}
  \begin{tabular}{p{3cm}p{2cm}p{2cm}}
    \toprule
    Reward condition        & POMDP  & MDP   \\ \midrule
    $r(o = o_{target})$     & 10000  & 10000 \\ 
    $r(v = \textrm{true})$  & -75    & -10   \\
    $r(n > n_{\max})$       & -75    & -10   \\
    otherwise $r(\cdot)$    & -100   & -1    \\
    \bottomrule
  \end{tabular}
\end{table}

The state observations are identical to the states that the mobile device can enter. 
In this case, however, uncertainty is introduced into the observation by the object detector.
Previous search locations and search time are fully observable here, since they can be tracked by the mobile device.
$\mathbf{\Omega}$ therefore only contains the classification/misclassification probabilities of the object detector. 
These values were found by performing a set of classifier tests on the object detector and generating a confusion matrix to populate $\mathbf{\Omega}$.

@@

\bibliographystyle{splncs04}
\bibliography{bib}

>>>>

\begin{abstract}
  The ActiVis project aims to deliver a mobile system that is able to guide a person with visual impairments towards a target object or area in an unknown indoor environment. 
  For this, it uses new developments in object detection, mobile computing, action generation and human-computer interfacing to interpret the user's surroundings and present effective guidance directions.
  Our approach to direction generation uses a Partially Observable Markov Decision Process (POMDP) to track the system's state and output the optimal location to be investigated.
  This system includes an object detector and an audio-based guidance interface to provide a complete active search pipeline.
  The ActiVis system was evaluated in a set of experiments showing better performance than a simpler unguided case.
  \keywords{Active vision, vision impairment, object detection}
\end{abstract}

@@

The issue with a marker-based approach is that it requires significant effort to place and maintain them in an environment, which is remedied by markerless systems.
However, both of these methods use passive guidance approaches that rely on the user placing the desired marker/object within the camera's view by themselves before any guidance is provided. 
VizWiz~\cite{bigham2010vizwiz} can leverage a human's understanding of the environment to guide a user to the correct location, but there is significant lag and a reliance on a good internet connection and remote worker being available. 
Previous work on the ActiVis system~\cite{lock2019active} addressed the passive guidance issue by implementing a Markov Decision Process (MDP) that gives the user pointing instructions to find an out-of-view object, showing that this is a viable method for object search. 
However, that work used QR-codes to simulate real objects and on-screen prompts to present the guidance instructions.
In this paper we expand the control model, replacing the QR-codes with real objects and providing audio guidance, thereby implementing a complete object detection and guidance pipeline for people with vision impairments.

@@

A complete system diagram for ActiVis is given in Fig.~\ref{fig:sys-diagram}. 
This diagram shows a typical feedback control loop that generates a control signal to minimise some error signal, $e$, and drive the output to some reference, $r$. In our case, the system incorporates a human within the loop; $r$ is the target object and $y$ is the actual observation of the mobile phone's camera.

@@

Adding a human into the loop means to have an additional block, $H$, representing a human that receives a control signal, $u$, from the controller, $K$. 
However, the challenge is that a person may interpret $u$ in some unpredictable way, resulting in the signal $u*$ that points the camera, $P$, to a new object observation,~$y$.
It is therefore important to design $K$ to be robust enough to accommodate different user habits and limitations, ensuring that $u*$ tracks $u$ as closely as possible. 
The object detector's classification errors, $n$, is added to the feedback loop as a noise signal that affects $K$'s output.

In our previous work~\cite{lock2019active}, we assumed a perfect object classifier and solved the problem using an MDP.
In this paper we remove that assumption and replace the MDP with a POMDP-based controller that can handle uncertainties in the object detection and classification. 
Our new controller works by generating a trail of virtual waypoints for the user to point the camera towards, which eventually leads to the target object.
The waypoint positions are based on the model's pre-trained internal knowledge of the inter-object spatial relationships and they are placed in a way that maximises the probability of the user finding the target object.
The design and implementation of this controller are discussed next.

@@

A POMDP is an extension of the MDP that handles cases where the state is not directly observable, allowing it to be used in more realistic scenarios. 
The implication of this is that a POMDP-based agent does not know its state at any point in time, but must infer it based on the known model parameters and sensor accuracy, which in our case is the object detector's accuracy.
This state inference relies on a so-called {\em belief meta-state}, which is updated with additional observations to reflect the likelihood that the agent is in any given state.
The belief state is fully observable by the agent and can be used to infer the mobile device's current state and generate the next action.

A POMDP model is described by the 8-tuple $(\mathbf{S}, \mathbf{A}, \mathbf{T}, \mathbf{R}, \mathbf{\Omega}, \mathbf{O}, \mathbf{b}, \gamma)$, where $\mathbf{S}$ represents a finite set of discrete states, $\mathbf{A}$ is a set of discrete actions, $\mathbf{T}$ is a matrix containing the probabilities of transitioning from state $s$ to state $s'$ (where $s, s' \in \mathbf{S}$) after executing action $a$ (where $a \in \mathbf{A}$), and $\mathbf{R}$ is the reward the agent receives executing action $a$ and reaching state $s'$.
$\mathbf{\Omega}$ is the set of possible state observations, while the matrix $\mathbf{O}$ contains the probabilities of making observation $\omega$ (where $\omega \in \mathbf{\Omega}$) when in state $s$ after executing action $a$. \tofix{sure about this? I think the observation is before the action...}
Finally, $\mathbf{b}$ is the belief vector containing the state probability distribution and $\gamma$ is a discount factor that prioritises long-term over short-term rewards, which affects the model's convergence rate. 
The POMDP's parameters and training process are discussed next.

@@

The state is given by $s = \langle u, n, v \rangle$, where $u\in\mathbf{U}$ is the object within view, $n$ the number of search steps taken during the search and $v$ is a binary variable that tracks the uniqueness of a waypoint's position for the search.

The possible actions that dictates the location where the mobile device will generate the next waypoint are given by $\mathbf{A} = \{ \textrm{UP, DOWN, LEFT, RIGHT} \}$.

$\mathbf{T}$ was determined by extracting the inter-object spatial relationships for a limited number of objects, in terms of the actions $\mathbf{A}$, from the OpenImages~\cite{openimages} dataset. 
For example, by iterating over the images containing the objects of interest, we can see that the object `monitor' is located above (i.e. UP) the object `keyboard' in 16\% of the images containing both objects (see Fig.~\ref{fig:openimage}). 
The transition function for this case is $t(s, a, s') = t(\textrm{keyboard}, \textrm{UP}, \textrm{monitor}) = 0.16$.

@@

The reward function encourages the device controller to search for the target object by giving a substantial reward if the object is found, while penalising the controller for every action that does not result in a successful object detection.
Furthermore, additional penalties are given if the controller generates a waypoint in an area it has explored before $(v = \textrm{true}$) or when it exceeds some search-length threshold denoted by $n_{\max}$.
The reward values were empirically determined starting from the implementation of our previous MDP model~\cite{lock2019active}, the rewards of which are summarised in Table~\ref{tab:rewards} together with the new ones. \tofix{can you explain why the last reward (otherwise) is much bigger than before, even bigger than the -75 rewards?}

\begin{table}
  \centering
  \caption{The reward function for the POMDP. }\label{tab:rewards}
  \begin{tabular}{p{3cm}p{2cm}p{2cm}}
    \toprule
    Reward condition        & POMDP  & MDP~\cite{lock2019active}   \\ \midrule
    $r(o = o_{target})$     & 10000  & 10000 \\ 
    $r(v = \textrm{true})$  & -75    & -10   \\
    $r(n > n_{\max})$       & -75    & -10   \\
    otherwise $r(\cdot)$    & -100   & -1    \\
    \bottomrule
  \end{tabular}
\end{table}

The state observations are identical to the states that the mobile device can enter into. 
In this case, however, uncertainty is introduced into the observation by the object detector.
Instead, the previous search locations and search time are fully observable, since they can be tracked by the mobile device.
$\mathbf{\Omega}$ therefore only contains the classification/misclassification probabilities of the object detector. 
These values were found by performing a set of classification tests with the object detector and generating a confusion matrix to populate $\mathbf{\Omega}$.

@@

\tofix{check all references to use correct information and full list of authors (suggest to use manual input on tools like JabRef)}
\bibliographystyle{splncs04}
\bibliography{bib}
